Generality in food webs scales with species richness, not latitude

### 2015-03-04 manuscript/latitudepaper.tex
Just changed which version of the figure to use
Added Tamara's comments

\section*{Introduction}

%Food webs are a thing. (Mention gen, vul, LS.) %Food webs have strong scaling structure.
Food webs --networks of feeding links between species-- have been used for over a century to summarize the structure of 
ecological communities \citep{Williams2000,earlierwork} and to understand how that structure relates to environmental variables 
such as land use \citep{Townsend1998,Townsend2004c,Digel2014}, primary productivity \citep{}, and climate\citep{}. The latter
variables in turn have strong gradients over latitude, with productivity and temperature both being higher in the tropics while 
climate is more variable at high latitudes \citep{}. These variables affect both the resources available and species' 
metabolisms~\citep{Stegen2012}, and have been proposed as influences on the strength of 
interspecific interactions~\citep{Marquis2005,Schemske2009}. 
By shaping interactions between species, latitudinal gradients may also shape food-web structure.
Indeed, these latitudinal environmental gradients have been put forward as potential driverd for the 
latitudinal gradient in species richness, one of the most general and robust patterns in ecology \citep{Schemske2009a,,}.

@@

One proposed link between species richness and latitude is the ``latitude-niche breadth hypothesis''~\citep{,Vazquez2004}. 
This hypothesis predicts that decreased seasonality in the tropics should lead to more stable populations, which in
turn should evolve smaller niches \citep{Vazquez2004}. These narrow niches should therefore allow more species to 
coexist in the tropics than can at higher latitudes. Alternatively, higher productivity of the tropics \citep{Brown2004}
may result in a broader niche space \citep{Davies2007} which could also allow greater diversification even if niche 
sizes are globally similar. Although the assumptions of the latitude-niche breadth hypothesis are only equivocably 
supported \citep{Vazquez2004}, it remains a compelling potential mechanism for the latitudinal gradient in species 
richness \citep{Lappalainen2006,Krasnov2008,Slove2010}.

If decreased seasonality or higher productivity is the cause of higher species-richness in the tropics, then 
there should be other effects of latitude on food-web structure. Attempts to unravel the effects, however,
are complicated by known effects of species richness on many other network properties~\citep{Riede2010}. In particular,
narrower niches imply fewer links per species (i.e., greater specialization in the tropics~\citep{}). However,
average numbers of links per species tend to increase in larger food webs~\citep{Dunne2006,Riede2010}. This means
that a latitudinal effect on specialization may be obscured by a latitudinal gradient in species richness. If this is
the case, it may still be possible to uncover effects of latitude on specialization by examining the shape of the scaling 
relationship between specialization and species richness over changing latitude. Here, we use three measures of specialization:
mean links per species, mean generality (number of prey), and mean vulnerability (number of predators). By testing whether
latitude affects the scaling of each property with species richness, we test effects of latitude on specialization implied by
the ``latitude-niche breadth hypothesis''.

@@

As the food webs in this dataset are derived from a variety of sources and were compiled over many decades, it
is likely that they vary in their resolution and in the amount of sampling effort invested in their assembly.
Many analyses of food-web structure reduce this variation by aggregating species with identical predator and prey
sets to form ``trophic species'' webs (e.g. ~\citep{Martinez1991,Vermaat2009,Dunne2004,Dunne2013}). As this study
is concerned directly with the number of species at a particular latitude, however, we did not wish to ignore 
species with redundant sets of interactions. We therefore analysed both original and trophic-species versions
of the dataset. As the version of the dataset used did not qualitatively change the results, we present only
the results for the original webs in the main text. In each case, we counted the number of species ($S$) and 
links in each web and used these to calculate the mean number of links per species ($Z$), mean generality 
($G$), and mean vulnerability ($V$) of each web.

@@

These effects varied strongly across ecosystem types. In marine and terrestrial food webs,
scaling of links per species, generality, and vulnerability was not strongly affected by
latitude. This is consistent with the idea that species' niche breadths do not vary systematically
with temperature and productivity but that the niche space might be larger in species-rich communities
~\citep{Davies2007}. In contrast, scaling in of lake, estuarine, and stream was generally stronger
towards the poles. This means that species in tropical communities gain fewer additional feeding links
per additional species in the web, supporting the hypothesis that greater stability in the tropics
leads to narrower niches~\citep{Brown2004} and a higher proportion of specialists~\citep{}. As food webs
are generally aggregated in space and time, the greater increase in links in high latitude stream, 
estuarine, and lake food webs might reflect prey switching as different food sources become
available~\citep{}.

@@

\begin{figure}[!h]
\label{scalingfig}
\includegraphics[width=.9\textwidth]{Figures/scaling_with_S_grey.eps}
\caption{Relationships between species richness and mean links per species, mean generality, and mean 
vulnerability (N=259 food webs). Scaling of mean links per species and mean vulnerability with species 
richness strengthened with latitude for estuarine, stream, and lake food webs. This strengthening was
reversed for marine and terrestrial food webs. Scaling of mean generality with species richness 
strengthened with latitude for lake food webs, but did not significantly vary with latitude for other
ecosystem types (N=259 food webs). }
\end{figure}

>>>>

\section*{Introduction} [[Add something about ecosystem type since it's in the methods]]

%Food webs are a thing. (Mention gen, vul, LS.) %Food webs have strong scaling structure.
Food webs --networks of feeding links between species-- have been used for over a century to summarize the structure of 
ecological communities \citep{Williams2000,earlierwork} and to understand how that structure relates to environmental variables 
such as land use [[lets really consider the three env variables used here. there really is not a lot of work on land use....i get why prim prod and climate are included....maybe focus on habitat types since you focus on habitats in the analysis]] \citep{Townsend1998,Townsend2004c,Digel2014}, primary productivity \citep{}, and climate\citep{}. The latter
variables in turn have strong gradients over latitude, with productivity and temperature both being higher in the tropics while 
climate is more variable at high latitudes \citep{}. [[maybe climatic regimes? the issue is that there really are not a lot of studies that look at these things.....]] These variables affect both the resources available and species' 
metabolisms~\citep{Stegen2012}, and have been proposed as determinants of the strength of 
interspecific interactions~\citep{Marquis2005,Schemske2009}. 
By modulating interactions between species, latitudinal gradients may also shape food-web structure.
Indeed, these latitudinal environmental gradients have been put forward as potential drivers for the 
latitudinal gradient in species richness, one of the most general and robust patterns in ecology \citep{Schemske2009a,,}.

@@

One proposed link between species richness and latitude is the ``latitude-niche breadth hypothesis''~\citep{,Vazquez2004}. 
This hypothesis predicts that decreased seasonality in the tropics should lead to more stable populations, which in
turn should evolve smaller niches \citep{Vazquez2004}. These narrow niches should therefore allow more species to 
coexist in the tropics than at higher latitudes. Alternatively, higher productivity of the tropics \citep{Brown2004}
may result in a broader niche space \citep{Davies2007} which could also allow greater diversification even if niche 
sizes are globally similar. Although the assumptions of the latitude-niche breadth hypothesis are only equivocably 
supported \citep{Vazquez2004}, it remains a compelling potential mechanism for the latitudinal gradient in species 
richness \citep{Lappalainen2006,Krasnov2008,Slove2010}.

If decreased seasonality or higher productivity is the cause of higher species-richness in the tropics, then 
there should be other effects of latitudinal gradients on food-web structure. Attempts to unravel these effects, however,
are complicated by known relationships between species richness and many other network properties~\citep{Riede2010}[[Check that I don't say that richness causes structural effects elsewhere]]. In particular,
narrower niches imply fewer links per species (i.e., greater specialization in the tropics~\citep{}). However,
average numbers of links per species tend to increase in larger food webs~\citep{Dunne2006,Riede2010}. This means
that a latitudinal effect on specialization may be obscured by a latitudinal gradient in species richness. If this is
the case, it may still be possible to uncover effects of latitude on specialization by examining the shape of the scaling 
relationship between specialization and species richness over changing latitude. Here, we use three measures of specialization:
mean links per species, mean generality (number of prey), and mean vulnerability (number of predators). By testing whether
latitude affects the scaling of each property with species richness, we test effects of latitude on specialization implied by
the ``latitude-niche breadth hypothesis''.

@@

As the food webs in this dataset are derived from a variety of sources and were compiled over many decades, it
is likely that they vary in their resolution and in the amount of sampling effort invested in their assembly.
Many analyses of food-web structure reduce this variation by aggregating species with identical predator and prey
sets to form ``trophic species'' webs (e.g. ~\citep{Martinez1991,Vermaat2009,Dunne2004,Dunne2013}). As this study
is concerned directly with the number of species at a particular latitude, however, we did not wish to ignore 
species with redundant sets of interactions. We therefore analysed both original and trophic-species versions
of the dataset. As the version of the dataset used did not qualitatively change the results, we present only
the results for the original webs in the main text. [[so what does this suggest? i find that in methods when making these statements it helps to go the step beyond and try and infer why trophic species/SR yielded similar results]] In each case, we counted the number of species ($S$) and 
links in each web and used these to calculate the mean number of links per species ($Z$), mean generality 
($G$), and mean vulnerability ($V$) of each web.

@@

These effects varied strongly across ecosystem types. In marine and terrestrial food webs,
scaling of links per species, generality, and vulnerability was not strongly affected by
latitude. This is consistent with the idea that species' niche breadths do not vary systematically
with temperature and productivity but that the niche space might be larger in species-rich communities
~\citep{Davies2007}. In contrast, scaling in of lake, estuarine, and stream was generally stronger
towards the poles. [[Sentence structure in following?]] This means that species in tropical communities gain fewer additional feeding links
per additional species in the web, supporting the hypothesis that greater stability in the tropics
leads to narrower niches~\citep{Brown2004} and a higher proportion of specialists~\citep{}. As food webs
are generally aggregated in space and time, the greater increase in links in high latitude stream, 
estuarine, and lake food webs might reflect prey switching as different food sources become
available~\citep{}.

@@

\begin{figure}[!h]
\label{scalingfig}
\includegraphics[width=.9\textwidth]{Figures/nonts_scaling_with_S_grey.eps}
\caption{Relationships between species richness and mean links per species, mean generality, and mean 
vulnerability (N=259 food webs). Scaling of mean links per species and mean vulnerability with species 
richness strengthened with latitude for estuarine, stream, and lake food webs. This strengthening was
reversed for marine and terrestrial food webs. Scaling of mean generality with species richness 
strengthened with latitude for lake food webs, but did not significantly vary with latitude for other
ecosystem types (N=259 food webs). }
\end{figure}
